Adaptive group-regularized logistic elastic net regression

### 2019-02-27 docs/manuscript.Rnw
changed the applications in manuscript.Rnw and supplement.Rnw

% settings
\graphicspath{{../figs/}}
\setlength{\evensidemargin}{.5cm} \setlength{\oddsidemargin}{.5cm}
\setlength{\textwidth}{15cm}
\title{Adaptive group-regularized logistic elastic net regression}
\date{}
\author{Magnus M. M\"unch$^{1,2}$\footnote{Correspondence to: 
\href{mailto:m.munch@vumc.nl}{m.munch@vumc.nl}}, Carel F.W. Peeters$^{1}$, 
Aad W. van der Vaart$^{2}$, and \\ Mark A. van de Wiel$^{1,3}$}

@@

	\section{Introduction}
	Prediction from high-dimensional data is a challenge common to many fields of
	research. Examples of such high-dimensional prediction problems arise in
	computer vision, stock market prediction and disease diagnosis from omics 
	data. In this paper, we specifically focus on the latter setting. High
	dimensionality of data introduces several issues in the estimation of 
	prediction models, especially: (a) unidentifiable models, (b) highly 
	correlated predictor variables and (c) non-trivial variable selection.

@@

	An example of such a study with extra information available is investigated in
	Section \ref{sec:app}. INTRODUCE DATA. This problem is not straightforward to
	solve with the standard methods available in literature. At first sight, such
	an \textit{a priori} grouping of the features suggests the group lasso
	\cite[]{meier_group_2008, yuan_model_2006} (or one of its extensions), which
	penalises and selects features at the group level. There are two limitations 
	to the group lasso: (a) it selects groups as a whole instead of single 
	features and (b) it is not adaptive: the amount of penalization a group
	receives depends on a `global' penalty parameter common to all groups.
	Extensions such as the sparse group lasso \cite[]{simon_sparse-group_2013}
	partly deal with (a), but do not address (b).

@@

	We introduce the Bayesian logistic (generalised) elastic net model in Section
	\ref{sec:model}. In Section \ref{seq:estimation} we derive a variational
	approximation to this model and use this novel approximation in the empirical
	Bayes estimation of multiple, group-specific penalty parameters. In Sections
	\ref{sec:simulations}, and \ref{sec:app}, we demonstrate the approach in a
	simulation study and a real data example, respectively. In Section 
	\ref{sec:discussion} we conclude by both discussing some differences and
	parallels between the proposed method, termed \texttt{gren}, and other methods
	in the literature.

@@

	Many forms of empirical Bayes exist, the most formal one being maximisation of
	the marginal likelihood with respect to the hyperparameters. The resulting 
	hyperparameter estimates are then plugged into the prior. The marginal 
	likelihood is often introduced as a measure of model evidence given the 
	observed data and is computed by integrating the product of likelihood and 
	prior with respect to the model parameters. In the case of the elastic net 
	introduced in (\ref{eq:enmodel2}) this results in the following marginal 
	empirical Bayes posterior for $\bbeta$:
	\begin{align}
	p_{\hat{\blambda}}(\bbeta | \y) & = \frac{\L (\y ; \bbeta) \pi_{\hat{\blambda}
	} (\bbeta) }{p_{\hat{\blambda}}(\y)} = \frac{\int_{\btau} \L (\y ; \bbeta) 
	\pi_{\hat{\blambda}}(\bbeta | \btau) \pi_{\hat{\blambda}}(\btau) \, d\btau}{p_
	{\hat{\blambda}}(\y)}, \nonumber \\
	\hat{\blambda} & := \underset{\blambda}{\argmax} \log p_{\blambda} (\y) = 
	\underset{\blambda}{\argmax} \int_{\bbeta} \int_{\btau} \L_{\blambda}(\y, 
	\bbeta, \btau) \, d\bbeta d\btau \nonumber \\
	& = \underset{\blambda}{\argmax} \int_{\bbeta} \int_{\btau} \L (\y ; \bbeta) 
	\pi_{\blambda}(\bbeta | \btau) \pi_{\blambda}(\btau) \, d\bbeta d\btau. 
	\label{eq:mmlestimate}
	\end{align}
	The integrals in (\ref{eq:mmlestimate}) are intractable in the case of the 
	elastic net. In the omics setting the integrals are also high-dimensional, in 
	which case numerical and Monte Carlo approximation methods become tedious and 
	computationally expensive. Moreover, Laplace approximation is known to suffer 
	from low accuracy in many high-dimensional settings 
	\cite[]{shun_laplace_1995}. In \cite{casella_empirical_2001} an EM algorithm 
	is described that estimates the hyperparameters. This EM algorithm iteratively
	maximises the expected joint log likelihood, such that the sequence:
	\begin{equation}\label{eq:ebEM}
	\blambda^{(k + 1)} = \underset{\blambda}{\argmax} \E_{\bbeta, \btau | \y} \[ 
	\log \L_{\blambda}(\y, \bbeta, \btau) | \blambda^{(k)} \]
	\end{equation}
	converges to a local maximum of the marginal likelihood. The difficulty herein
	is in the calculation of the expected joint log likelihood. 
	\cite{casella_empirical_2001} suggests to approximate the expectation by its Monte Carlo expectation. Although elegant and simple, this method requires a converged MCMC sample from the posterior for every iteration: a computationally intensive procedure. \cite{roy_selection_2017} introduce generalised importance sampling for the Bayesian elastic net, such that we just need a limited, pre-specified number of MCMC chains. However, this still requires several converged chains, so is not feasible in many high-dimensional omics settings. We propose to tackle this problem by approximating the expectation in (\ref{eq:ebEM}) using variational Bayes.

@@

	After a change of variables $\psi_j = \tau_j - 1$, we find the optimal distributions in our variational Bayes implementation for the model parameters as:
	\begin{equation}\label{eq:varbayesdistr}
	q^*_{\bbeta} (\bbeta) \sim \mathcal{N} (\bmu, \bSigma) \text{, } q^*_{\bomega} (\bomega) \sim \prod_{i=1}^n \mathcal{PG} (m_i, c_i)\text{, and } q^*_{\bpsi}(\bpsi) \sim \prod_{j=1}^p \mathcal{GIG} \(\frac{1}{2}, \frac{\lambda_1^2}{4 \lambda_2}, \chi_j\),
	\end{equation}
	where $\mathcal{GIG} (\cdot)$ denotes the generalized inverse Gaussian distribution (See SM Section 5 for the derivations). The so-called variational parameters in (\ref{eq:varbayesdistr}) contain cyclic dependencies, so we update them by:
	\begin{subequations}\label{eq:VBupdateequations}
		\begin{align}
		\bSigma^{(t + 1)} &= \( \X \tr \bOmega^{(t)} \X + \lambda_2 \bLambda' + \frac{\lambda_1 \sqrt{\lambda_2}}{2} \bLambda' \Z^{(t)}\)^{-1}, \\ & \text{with } \bOmega^{(t)} = \diag\[ \(\frac{m_i}{2 c_i^{(t)}}\) \tanh \( \frac{c_i^{(t)}}{2} \)\] \text{ and } \Z^{(t)}=\diag\[(\chi_j^{(t)})^{-1/2}\], \nonumber \\
		\bmu^{(t + 1)} &= \bSigma^{(t + 1)} \X \tr (\y - \m/2), \\
		c_i^{(t + 1)} &= \sqrt{\x_i \tr \bSigma^{(t + 1)} \x_i + (\x_i \tr \bmu^{(t + 1)})^2} \text{, for } i=1, \dots, n, \\
		\chi_j^{(t + 1)} &= \lambda'_{g(j)} \lambda_2 \[\bSigma^{(t + 1)}_{jj} + (\bmu^{(t + 1)}_j)^2 \] \text{, for } j=1, \dots, p,
		\end{align}
	\end{subequations}
	until convergence. Here, $\bLambda'$ is a diagonal matrix with entries $\lambda'_g$, each repeated $|\G(g)|$ times and $\m = \begin{bmatrix} m_1 & \cdots & m_n \end{bmatrix} \tr$. Furthermore, $\mathbf{A}_{jj}$ and $\mathbf{a}_j$ denote the $j$th diagonal element of a square matrix and $j$th element of a column vector, respectively. Naive calculation of the variational parameters is computationally expensive. In Section 6 of the SM we show that informed calculation of the parameters results in a significant reduction of computational complexity.

	\subsection{Empirical-variational Bayes}\label{seq:empvarbayes}
	Variational Bayes was shown to underestimate the posterior variance of the parameters, both numerically and theoretically, in several settings \cite[]{rue_approximate_2009,consonni_mean-field_2007,bishop_pattern_2006,wang_inadequacy_2005}. This coincides with our experience that the global penalty parameters $\lambda_1$ and $\lambda_2$ tend to be overestimated. To prevent overestimation we use the parametrisation of the elastic net in \cite{friedman_regularization_2010} as introduced in Section \ref{seq:empiricalbayes}: we fix $\alpha$ and estimate $\lambda$ by cross-validation of the regular elastic net model, such that the overall penalisation is determined by cross-validation of only $\lambda$. By combining cross-validation of the global penalty parameter $\lambda$ with empirical Bayes estimation of the penalty multipliers $\blambda' = \begin{bmatrix} \lambda'_1 & \cdots & \lambda'_G \end{bmatrix} \tr$, the estimation is more robust to underestimation of the variational posterior variances. The remaining issue is the choice of $\alpha$. \cite{hastie_glmnet_2016} recommend to either fix $\alpha$ \textit{a priori}, or compare the results for several choices of $\alpha$. We recommend the latter. Otherwise we note that, in our experience, the choice $\alpha=0.5$ often gives good results.

@@

	\section{Simulations}\label{sec:simulations}
	\subsection{Setup}\label{sec:setup}
	To assess variable selection and predictive performance we conduct a simulation study in which we compare \texttt{gren} to the regular elastic net and ridge models, \texttt{GRridge} \cite[]{van_de_wiel_better_2016}, sparse group lasso (\texttt{SGL}) \cite[]{simon_sparse-group_2013}, group mimimax concave penalty (\texttt{grMCP}) \cite[]{breheny_penalized_2009}, and the group exponential lasso (\texttt{gel}) \cite[]{breheny_group_2015}. \texttt{GRridge} is similar to \texttt{gren} in the sense that it estimates group-specific penalty multipliers. The two main differences with \texttt{gren} are (i) the absence of an $L_1$-norm penalty and (ii) the estimation procedure. The other methods are extensions of the group lasso \cite[]{meier_group_2008, yuan_model_2006} and not adaptive on the group level. In contrast to the original group lasso, all of them are able to select single features, instead of just on the group level.

	For the regular elastic net, \texttt{gren}, \texttt{SGL}, \texttt{grMCP}, and \texttt{gel}, we fix the proportion of $L_1$-norm penalty to three different values: $\alpha \in \{0.05, 0.5, 0.95 \}$. The first setting closely resembles the ridge setting, where $\alpha=0$, while the third setting is similar to the lasso with $\alpha=1$. For the methods that do automatic variable selection (all methods except the regular ridge and \texttt{GRridge}), we estimate the models for a range of model sizes.

	Predictive performance is measured by area under the receiver operator curve (AUC) and Brier skill score (BSS). AUC gives the area under the curve of sensitivity versus $1-$ specificity and equals the probability that the classifier ranks a randomly chosen case higher than a randomly chosen non-case. The BSS is a normalised version of the Brier score: $\text{BSS} = 1 - \frac{\sum_{i=1}^n (y_i - \hat{y}_i)^2}{\sum_{i=1}^n (y_i - \bar{y})^2}$,
	where $\hat{y}_i=\expit(\x_i \tr \hat{\bbeta})$ is the prediction for observation $i$ and $\bar{y}=n^{-1} \sum_{i=1}^n y_i$. The BSS measures the fraction of Brier score improvement compared to a non-informative model. Quality of feature selection is assessed by Cohen's kappa \cite[]{cohen_coefficient_1960} and parameter estimation by mean squared error (MSE). Both are presented in SM Section 10.

	We simulated data according to five different scenarios:
	\begin{enumerate}
		\item differential signal between the groups and uniformly distributed model parameters,
		\item a large number of small groups of features,
		\item no differential signal between the groups, but strong correlations within groups of features,
		\item differential signal between the groups and heavy-tailed distributed model parameters, and
		\item a very sparse setting with no signal in some of the groups.
	\end{enumerate}
	Full descriptions of the scenarios are given in Section 10 of the SM. In the first scenario, the distribution of the model parameters is far off from the assumed ridge/elastic net/lasso distributions, so we assume that all methods will underperform. In the second scenario, we expect \texttt{gren} and \texttt{GRridge} to overfit due to the large number of hyperparameters to estimate. The rest of the methods are conjectured to perform better, due to a lack of overfitting. In scenario three, we expect all methods to perform badly due to the high correlation between the features. In scenario three

	In all scenarios we created a training set of $n=100$ samples to estimate the models and a test set of $n_{\text{test}}=1000$ samples on which we compute the performance measures. The features $\x_i$, $i=1, \dots, n$, are sampled from a $p$-dimensional Gaussian distribution centred at zero, where we set $p=1000$. We let the covariance matrix of the features vary over the scenarios. Likewise for the model parameters $\beta_j$, $j=1, \dots, p$. We simulate the outcome data $y_i$, $i=1, \dots, n$, from a binary logistic model: $y_i \sim \mathcal{B}(1, \expit (\x_i \tr \bbeta))$. To mitigate the influence of random variation in the simulations, we repeated every scenario 100 times and report estimated penalty parameters and LOESS smoothed averages of the performance measures for a range of model sizes.

	\subsection{Results}\label{sec:results}

	\section{Application to metabolomics in alzheimer}\label{sec:app}
	\subsection{Co-data separately included}
<<lines_metabolomics_alzheimer_res1_auc, fig.cap="Cross-validated AUC in alzheimer example using metabolomics class co-data, with a) the full range and b) cut-out", out.width="100%", fig.asp=2/3>>=
@

<<barplots_metabolomics_alzheimer_res1_auc, fig.cap="Estimated penalty multipliers in alzheimer example using metabolomics class co-data", out.width="50%", fig.asp=1>>=
@

<<lines_metabolomics_alzheimer_res2_auc, fig.cap="Cross-validated AUC in alzheimer example using quality score co-data, with a) the full range and b) cut-out", out.width="100%", fig.asp=2/3>>=
@

<<barplots_metabolomics_alzheimer_res2_auc, fig.cap="Estimated penalty multipliers in alzheimer example using quality score co-data", out.width="50%", fig.asp=1>>=
@

<<lines_metabolomics_alzheimer_res3_auc, fig.cap="Cross-validated AUC in alzheimer example using node degree co-data, with a) the full range and b) cut-out", out.width="100%", fig.asp=2/3>>=
@

<<barplots_metabolomics_alzheimer_res3_auc, fig.cap="Estimated penalty multipliers in alzheimer example using node degree co-data", out.width="50%", fig.asp=1>>=
@

\subsection{Co-data included simultaneously}
<<lines_metabolomics_alzheimer_res4_auc, fig.cap="Cross-validated AUC in alzheimer example using all co-data, with a) the full range and b) cut-out", out.width="100%", fig.asp=2/3>>=
@

<<barplots_metabolomics_alzheimer_res4_auc, fig.cap="Estimated penalty multipliers in alzheimer example using all co-data for a) metabolomics class, b) quality score, and c) node degree", out.width="100%", fig.asp=1>>=
@

>>>>

% settings
\graphicspath{{../figs/}}
\setlength{\evensidemargin}{.5cm} \setlength{\oddsidemargin}{.5cm}
\setlength{\textwidth}{15cm}
\title{Adaptive group-regularized logistic elastic net regression}
\date{\today}
\author{Magnus M. M\"unch$^{1,2}$\footnote{Correspondence to: 
\href{mailto:m.munch@vumc.nl}{m.munch@vumc.nl}}, Carel F.W. Peeters$^{1}$, 
Aad W. van der Vaart$^{2}$, and \\ Mark A. van de Wiel$^{1,3}$}

@@

	\section{Introduction}\label{sec:introduction}
	Prediction from high-dimensional data is a challenge common to many fields of
	research. Examples of such high-dimensional prediction problems arise in
	computer vision, stock market prediction and disease diagnosis from omics 
	data. In this paper, we specifically focus on the latter setting. High
	dimensionality of data introduces several issues in the estimation of 
	prediction models, especially: (a) unidentifiable models, (b) highly 
	correlated predictor variables and (c) non-trivial variable selection.

@@

	An example of such a study with extra information available is investigated in
	Sections [! section refs here]. The first one is concerned with treatment 
	response prediction for colorectal cancer patients, based on sequenced
	microRNAs. In the second example we attempt to predict lymph node metastases,
	using RNAseq data. In both settings we have additional information on the 
	features in the form of a partitioning available. Furthermore, in both
	examples data is sparse, so we would like to 
	include the additional information to enhance predictive performance of our
	models. Inclusion of the partitions is not straightforward using standard 
	methods available in literature. At first sight, such
	an \textit{a priori} grouping of the features suggests the group lasso
	\cite[]{meier_group_2008, yuan_model_2006} (or one of its extensions), which
	penalises and selects features at the group level. There are two limitations 
	to the group lasso: (a) it selects groups as a whole instead of single 
	features and (b) it is not adaptive: the amount of penalization a group
	receives depends on a `global' penalty parameter common to all groups.
	Extensions such as the sparse group lasso \cite[]{simon_sparse-group_2013}
	partly deal with (a), but do not address (b).

@@

	We introduce the Bayesian logistic (generalised) elastic net model in Section
	\ref{sec:model}. In Section \ref{seq:estimation} we derive a variational
	approximation to this model and use this novel approximation in the empirical
	Bayes estimation of multiple, group-specific penalty parameters. In Section
	\ref{sec:simulations} and Sections [! section ref here] and 
	[! section ref here], we demonstrate the approach in a
	simulation study and a two real data examples, respectively. In Section 
	\ref{sec:discussion} we conclude by both discussing some differences and
	parallels between the proposed method, termed \texttt{gren}, and other methods
	in the literature.

@@

	Many forms of empirical Bayes exist, the most formal one being maximisation of
	the marginal likelihood with respect to the hyperparameters. The resulting 
	hyperparameter estimates are then plugged into the prior. The marginal 
	likelihood is often introduced as a measure of model evidence given the 
	observed data and is computed by integrating the product of likelihood and 
	prior with respect to the model parameters. In the case of the elastic net 
	introduced in (\ref{eq:enmodel2}) this results in the following marginal 
	empirical Bayes posterior for $\bbeta$:
	\begin{align}
	  p_{\hat{\blambda}}(\bbeta | \y) & = \frac{\L (\y ; \bbeta) 
	  \pi_{\hat{\blambda}} (\bbeta) }{p_{\hat{\blambda}}(\y)} = 
	  \frac{\int_{\btau} \L (\y ; \bbeta) 
	  \pi_{\hat{\blambda}}(\bbeta | \btau) \pi_{\hat{\blambda}}(\btau) \,
	  d\btau}{p_{\hat{\blambda}}(\y)}, \nonumber \\
	  \hat{\blambda} & := \underset{\blambda}{\argmax} \log p_{\blambda} (\y) = 
	  \underset{\blambda}{\argmax} \int_{\bbeta} \int_{\btau} \L_{\blambda}(\y, 
	  \bbeta, \btau) \, d\bbeta d\btau \nonumber \\
	  & = \underset{\blambda}{\argmax} \int_{\bbeta} \int_{\btau} \L (\y ; \bbeta)
	  \pi_{\blambda}(\bbeta | \btau) \pi_{\blambda}(\btau) \, d\bbeta d\btau. 
	  \label{eq:mmlestimate}
	\end{align}
	The integrals in (\ref{eq:mmlestimate}) are intractable in the case of the 
	elastic net. In the omics setting the integrals are also high-dimensional, in 
	which case numerical and Monte Carlo approximation methods become tedious and 
	computationally expensive. Moreover, Laplace approximation is known to suffer 
	from low accuracy in many high-dimensional settings 
	\cite[]{shun_laplace_1995}. In \cite{casella_empirical_2001} an EM algorithm 
	is described that estimates the hyperparameters. This EM algorithm iteratively
	maximises the expected joint log likelihood, such that the sequence:
	\begin{equation}\label{eq:ebEM}
	\blambda^{(k + 1)} = \underset{\blambda}{\argmax} \E_{\bbeta, \btau | \y} \[ 
	\log \L_{\blambda}(\y, \bbeta, \btau) | \blambda^{(k)} \]
	\end{equation}
	converges to a local maximum of the marginal likelihood. The difficulty herein
	is in the calculation of the expected joint log likelihood. 
	\cite{casella_empirical_2001} suggests to approximate the expectation by its 
	Monte Carlo expectation. Although elegant and simple, this method requires a 
	converged MCMC sample from the posterior for every iteration: a 
	computationally intensive procedure. \cite{roy_selection_2017} introduce
	generalised importance sampling for the Bayesian elastic net, such that we
	just need a limited, pre-specified number of MCMC chains. However, this still
	requires several converged chains, so is not feasible in many high-dimensional
	omics settings. We propose to tackle this problem by approximating the
	expectation in (\ref{eq:ebEM}) using variational Bayes.

@@

	After a change of variables $\psi_j = \tau_j - 1$, we find the optimal 
	distributions in our variational Bayes implementation for the model parameters
	as:
	\begin{equation}\label{eq:varbayesdistr}
	  q^*_{\bbeta} (\bbeta) \sim \mathcal{N} (\bmu, \bSigma) \text{, } 
	  q^*_{\bomega} (\bomega) \sim \prod_{i=1}^n \mathcal{PG} (m_i, c_i)
	  \text{, and } q^*_{\bpsi}(\bpsi) \sim \prod_{j=1}^p \mathcal{GIG} 
	  \(\frac{1}{2}, \frac{\lambda_1^2}{4 \lambda_2}, \chi_j\),
	\end{equation}
	where $\mathcal{GIG} (\cdot)$ denotes the generalized inverse Gaussian 
	distribution (See SM Section 5 for the derivations). The so-called variational
	parameters in (\ref{eq:varbayesdistr}) contain cyclic dependencies, so we 
	update them by:
	\begin{subequations}\label{eq:VBupdateequations}
		\begin{align}
		  \bSigma^{(t + 1)} &= \( \X \tr \bOmega^{(t)} \X + \lambda_2 \bLambda' + 
		  \frac{\lambda_1 \sqrt{\lambda_2}}{2} \bLambda' \Z^{(t)}\)^{-1}, \\ 
		  & \text{with } \bOmega^{(t)} = \diag\[ \(\frac{m_i}{2 c_i^{(t)}}\) \tanh 
		  \( \frac{c_i^{(t)}}{2} \)\] \text{ and } \Z^{(t)}=
		  \diag\[(\chi_j^{(t)})^{-1/2}\], \nonumber \\
		  \bmu^{(t + 1)} &= \bSigma^{(t + 1)} \X \tr (\y - \m/2), \\
		  c_i^{(t + 1)} &= \sqrt{\x_i \tr \bSigma^{(t + 1)} \x_i + 
		  (\x_i \tr \bmu^{(t + 1)})^2} \text{, for } i=1, \dots, n, \\
		  \chi_j^{(t + 1)} &= \lambda'_{g(j)} \lambda_2 \[\bSigma^{(t + 1)}_{jj} + 
		  (\bmu^{(t + 1)}_j)^2 \] \text{, for } j=1, \dots, p,
		\end{align}
	\end{subequations}
	until convergence. Here, $\bLambda'$ is a diagonal matrix with entries 
	$\lambda'_g$, each repeated $|\G(g)|$ times and $\m = \begin{bmatrix} m_1 & 
	\cdots & m_n \end{bmatrix} \tr$. Furthermore, $\mathbf{A}_{jj}$ and 
	$\mathbf{a}_j$ denote the $j$th diagonal element of a square matrix and $j$th
	element of a column vector, respectively. Naive calculation of the variational
	parameters is computationally expensive. In Section 6 of the SM we show that 
	informed calculation of the parameters results in a significant reduction of 
	computational complexity.

	\subsection{Empirical-variational Bayes}\label{seq:empvarbayes}
	Variational Bayes was shown to underestimate the posterior variance of the 
	parameters, both numerically and theoretically, in several settings 
	\cite[]{rue_approximate_2009,consonni_mean-field_2007,bishop_pattern_2006,
	wang_inadequacy_2005}. This coincides with our experience that the global 
	penalty parameters $\lambda_1$ and $\lambda_2$ tend to be overestimated. To 
	prevent overestimation we use the parametrisation of the elastic net in 
	\cite{friedman_regularization_2010} as introduced in Section 
	\ref{seq:empiricalbayes}: we fix $\alpha$ and estimate $\lambda$ by 
	cross-validation of the regular elastic net model, such that the overall 
	penalisation is determined by cross-validation of only $\lambda$. By combining
	cross-validation of the global penalty parameter $\lambda$ with empirical 
	Bayes estimation of the penalty multipliers $\blambda' = \begin{bmatrix} 
	\lambda'_1 & \cdots & \lambda'_G \end{bmatrix} \tr$, the estimation is more 
	robust to underestimation of the variational posterior variances. The 
	remaining issue is the choice of $\alpha$. \cite{hastie_glmnet_2016} recommend
	to either fix $\alpha$ \textit{a priori}, or compare the results for several 
	choices of $\alpha$. We recommend the latter. Otherwise we note that, in our 
	experience, the choice $\alpha=0.5$ often gives good results.

@@

	\section{Simulations}\label{sec:simulations}
	To assess variable selection and predictive performance we conducted a 
	simulation study in which we compare \texttt{gren} to the regular elastic 
	net and ridge models, \texttt{GRridge} \cite[]{van_de_wiel_better_2016}, 
	sparse group lasso (\texttt{SGL}) \cite[]{simon_sparse-group_2013}, 
	group mimimax concave penalty (\texttt{grMCP}) 
	\cite[]{breheny_penalized_2009}, and the group exponential lasso 
	(\texttt{gel}) \cite[]{breheny_group_2015}. \texttt{GRridge} is similar to 
	\texttt{gren} in the sense that it estimates group-specific penalty 
	multipliers. The two main differences with \texttt{gren} are (i) the absence 
	of an $L_1$-norm penalty and (ii) the estimation procedure. The other methods 
	are extensions of the group lasso \cite[]{meier_group_2008, yuan_model_2006} 
	and not adaptive on the group level. In contrast to the original group lasso,
	all of them are able to select single features, instead of just on the group 
	level.

	For the regular elastic net, \texttt{gren}, \texttt{SGL}, \texttt{grMCP}, 
	and \texttt{gel}, we fix $\alpha \in \{0.05, 0.5, 0.95 \}$. In \texttt{gren}
	and the elastic net, $\alpha$ determines the proportion of $L_1$-norm penalty,
	such that the first setting closely resembles the ridge setting, 
	where $\alpha=0$, while the third setting is similar to the lasso with 
	$\alpha=1$. In \texttt{SGL}, \texttt{grMCP}, and \texttt{gel}, $\alpha$
	determines the proportion of group-level penalization ($\alpha=0$) compared to
	$L_1$-norm penalty on the feature level ($\alpha=1$). For the methods that do 
	automatic variable selection, we estimate 
	the models for a range of model sizes.

	We simulated data according to five different scenarios:
	\begin{enumerate}
		\item differential signal between the groups and uniformly distributed model
		parameters,
		\item a large number of small groups of features,
		\item no differential signal between the groups, but strong correlations 
		within groups of features,
		\item differential signal between the groups and heavy-tailed distributed 
		model parameters, and
		\item a very sparse setting with no signal in some of the groups.
	\end{enumerate}
	Full descriptions of the scenarios are given in Section [! ref to section] of 
	the SM. In the first scenario, the distribution of the model parameters is far
	off from the assumed ridge/elastic net/lasso distributions, so we assume that 
	all methods will underperform. In the second scenario, we expect \texttt{gren}
	and \texttt{GRridge} to overfit due to the large number of hyperparameters to 
	estimate. The rest of the methods are conjectured to perform better, due to a
	lack of overfitting. In scenario three, we expect all methods to perform badly
	due to the high correlation between the features. The \texttt{gren} fourth 
	scenario is conjectured to give a better \texttt{gren} performance, due to the
	correctly assumed heavier-tailed distribution and differential signal between
	the groups. Scenario five is expected to lead to a bad performance in all 
	models, because of the sparseness of the truth and lack of differential 
	signal.

	In all scenarios we created a training set of $n=100$ samples to estimate the 
	models and a test set of $n_{\text{test}}=1000$ samples on which we compute 
	the performance measures. The features $\x_i$, $i=1, \dots, n$, are taken 
	from a $p$-dimensional zero-centred Gaussian distribution, where we set 
	$p=1000$. We let the covariance matrix of the features vary over the 
	scenarios. Likewise for the model parameters $\beta_j$, $j=1, \dots, p$. We 
	simulate the outcomes $y_i$, $i=1, \dots, n$, from a binary logistic model: 
	$y_i \sim \mathcal{B}(1, \expit (\x_i \tr \bbeta))$. To mitigate the influence
	of random variation in the simulations, we repeated every scenario 100 times 
	and report estimated penalty parameters and predictive performance measures 
	area under the receiver operator curve (AUC) and Brier skill score (BSS).

	[! results here]

	Additional results are presented in Section [! section ref here] of the SM.

	\section{Application to microRNAs in colorectal cancer}\label{sec:appcolon}
  The presented data is from a deep sequencing analysis on microRNAs 
  \cite[]{neerincx_combination_2018}. The study was done in 88 treated 
  colorectal cancer patients, with the aim of predicting treatment response,
  coded as either non-progressive/remission (70 patients) or progressive 
  (18 patients). After pre-processing and normalisation, 2114 microRNAs 
  remained. In addition to the 2114 microRNAs, four clinical covariates were
  included in the analysis: prior use of adjuvant therapy (binary),
  the type of systemic treatment regimen (ternary), age, and primary tumor
  differentiation (binary). These clinical covariates are included as 
  unpenalized features in the model.

  In a preliminary experiment on different subjects, the microRNA expression 
  levels of metastatic colorectal tumour tissue were compared to normal 
  non-colorectal tissue and primary colorectal tumour tissue was compared to
  primary colorectal tumour tissue \cite[]{neerincx_mir_2015}. This yielded 221
  microRNAs that were differentially expressed in both comparisons ($\text{FDR}
  \leq 0.05$), versus 1893 not differentially expressed microRNAs. We expect
  that incorporation of this partitioning enhances therapy response 
  classification, because tumor-specific miRNAs might be more relevant than 
  non-specific ones. In addition, we divided the differentially expressed
  microRNAs even further into 127 highly differentially expressed microRNAs
  ($\text{FDR} \leq 0.001$) and 94 medium differentially expressed microRNAs 
  ($0.001 < \text{FDR} \leq 0.05$).  We will refer to this second partitioning
  as the three-group setting, as opposed to the first, the two-group setting.

  We split the data into 61 training, and 27 test instances and
  compared the performance of \texttt{gren} on the test set to ridge, 
  \texttt{GRridge}, elastic net, and the composite minimax
  concave penalty (\texttt{cMCP}) by \cite{breheny_penalized_2009}. Further
  comparisons to the sparse
  group lasso (\texttt{SGL}) by \cite{simon_sparse-group_2013} and the group
  exponential lasso (\texttt{gel}) \cite[]{breheny_group_2015} are presented in 
  the SM [! section ref here]. We compared the methods on AUC. More performance
  measures are presented in the SM [! section ref here].
  For the elastic net methods (including \texttt{gren}) and \texttt{cMCP}, we 
  set $\alpha \in \{ 0.05, 0.5, 0.95\}$. We experienced numerical issues with 
  \texttt{GRridge} when cross-validating 
  the global penalty parameter on the training data. We therefore estimated this
  global parameter on the full data for \texttt{GRridge}.

  The results are presented in Figure 
  \ref{fig:barplot_lines_micrornaseq_colorectal_cancer_res2_auc}. 
  The penalty multipliers estimated by 
  \texttt{gren} and \texttt{GRridge} are according to expectation: the highly
  expressed group receives the lowest penalty, followed by the medium expressed
  group, while the non-expressed group receives the strongest penalty. 
  \texttt{gren} seems to outperform the other methods for a range of model 
  sizes. This is especially pronounced for the larger models. The performance of
  \texttt{cmcp} is somewhat unstable. We noticed this is in general
  for the group-lasso and its extensions (see SM [! section ref here]). Overall,
  it seems as though the inclusion of the extra information on the features 
  benefited predictive performance: both \texttt{GRridge} and \texttt{gren} 
  outperform their respective non-group-regularized versions.
<<barplot_lines_micrornaseq_colorectal_cancer_res2_auc, fig.cap="Results for the colorectal cancer example with (a) the estimated penalty multipliers and (b) estimated AUC.", out.width="100%", fig.asp=1/2>>=
@

  \section{Application to RNAseq in lymph node cancer}\label{sec:applymph}
  In this application, the aim is to predict lymph node metastasis in patients 
  with HPV negative oral cancer using sequenced RNAs from TCGA 
  \cite[]{the_cancer_genome_atlas_network_comprehensive_2015}. The features are 
  normalised TCGA RNASeqv2 profiles of head-and-neck squamous cell 
  carcinomas on 133 patients with HPV-negative tumours in the oral cavity. Of 
  these patients, 76 had lymph node metastasis, and 57 did not. For a thorough 
  introduction of these data, see \cite{te_beest_improved_2017}.

  As co-data we considered two sources: (a) the cis-correlation between TCGA DNA
  copy number measurements and the RNASeqv2 data on the same patients, 
  quantified through Kendall's $\tau$ and binned into five groups, and (b) 
  p-values from an external microarray data set \cite[]{mes_prognostic_2017}, 
  binned into five groups. We expect features 
  with a positive Kendall's $\tau$ to be more important in metastasis prediction   
  \cite[]{masayesva_gene_2004}. Likewise, we expect features with a low 
  p-value to be more important.

  We split our data into 120 training instances and 97 test instances, trained
  the models on the training set and estimated AUC on the test set. Other 
  performance measures are presented in the SM Section [! section ref here]. We
  compared the same models as in Section \ref{sec:appcolon}. In this 
  application, the co-data consists of two partitions with overlapping groups.
  \texttt{cMCP} is not designed to deal with overlapping groups, so we 
  cross-tabulate the two co-data partitions to create one paritioning with 25 
  groups for \texttt{cMCP}. In addition, we compare to an extension of 
  \texttt{cMCP} that is able to handle multiple groupings 
  \cite[]{zeng_overlapping_2016}. We present the results in Figure 
  \ref{fig:barplot_lines_rnaseq_lymph_node_metastasis_res1_auc}. 
  Additional comparisons to \texttt{SGL}, \texttt{gel} and 
  the overlapping group version of \texttt{gel} \cite[]{zeng_overlapping_2016}
  \cite[]{breheny_group_2015} are found in the SM Section [! section ref here].
<<barplot_lines_rnaseq_lymph_node_metastasis_res1_auc, fig.cap="Results for the lymph node metastasis example with (a) the estimated penalty multipliers for the cis-correlations, (b) the estimated penalty multipliers for the p-values, and (c) the estimated AUC.", out.width="100%", fig.asp=1>>=
@

  We present additional applications and comparisons in the SM Sections 
  [! sections ref here].
